Cluster Difference Imaging Photometric Survey (CDIPS) II: Validatation of a Jupiter-Sized Planet in IC~2602

### 2020-06-18 paper/ms.tex
restructure paper

\begin{abstract}
  We present the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA {\it Transiting Exoplanet Survey Satellite} (TESS), the ESA
  Gaia mission, ground-based photometry, and spectroscopy from
  CTIO1.5/CHIRON, MPG2.2/FEROS, and AAT/Veloce.  We find that
  TOI$\,$837 is a $T=9.9$ G-dwarf in IC~2602 (the ``Southern
  Pleiades'').  The star and validated planet are therefore 30$\pm$20
  million years old.  The planet is warm ($P = 8.YY\,{\rm d}$) and
  roughly Jupiter-sized.  However, its transits are grazing: $b >
  0.92$ at 3$\sigma$.  From TESS photometry alone, the planetary size
  lies within 0.58--3.87 $R_{\rm Jup}$, due to the degeneracy between
  the planetary size and the impact parameter of the transit.  From
  radial velocity monitoring, we limit the mass of planet to less than
  2.XX $M_{\rm Jup}$ (3$\sigma$).  Grazing transits are cause for
  concern, as they usually are indicative of astrophysical false
  positive scenarios.  Our follow-up data render such scenarios highly
  unlikely.  Multi-color photometry requires eclipsing companions to
  have nearly identical color as TOI$\,$837, eliminating hierarchical
  eclipsing binary scenarios.  Background eclipsing binary scenarios
  are limited by speckle imaging, but formally remain a 0.2\%
  possibility.  TOI 837.01 therefore joins the ranks of adolescent
  exoplanets orbiting bright host stars as a validated planet.
  Further analysis of its mass, atmosphere, and obliquity should
  confirm its planetary nature beyond doubt, and are worthwhile
  because young planets are expected to do interesting things.
\end{abstract}

@@

Section~\ref{sec:observations} 
Section~\ref{sec:star}
Section~\ref{sec:planet}
Section~\ref{sec:discussion}
Section~\ref{sec:conclusions}.

\section{The Data}
\label{sec:observations}

@@

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Phase-folded transit of \pn.} Gray points are
    2-minute PDCSAP flux measurements, detrended with
    a 0.3-day robust Huber spline (see Section~\ref{subsec:tess}).
    Black points are binned to approximately 1 point per 6 minutes.
    The blue line shows the transit model corresponding to the median
    transit parameters (Table~2).
    % The bottom panel shows the residuals.
    \label{fig:phasefold}
	}
\end{figure*}

@@

\section{The Star}
\label{sec:star}

\subsection{Cluster Membership}
\label{subsec:member}

\subsubsection{HR Diagram}
\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ 
  {\bf Hertzsprung-Russell diagram of \tn\ and members of \cn.}
  Members (black circles) were identified by
  \citet{cantatgaudin_gaia_2018}.  Gray circles are non-member stars
  within 5 standard deviations of the mean \cn\ right ascension,
  declination, and parallax.  $G$ denotes Gaia broadband magnitudes,
  $Bp$ Gaia blue, $Rp$ Gaia red, and $\omega_{\rm as}$ the parallax in
  arcseconds. 
  \label{fig:hr}
	}
\end{figure*}

\begin{figure}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f9.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Rotation periods for \tn\ and selected open clusters.}
    The Pleiades (120$\,$Myr), Praesepe (670$\,$Myr), and NGC$\,$6811
    (1000$\,$Myr) are shown.  Their rotation periods were measured by
    \citet{rebull_rotation_2016a,douglas_poking_2017,douglas_k2_2019},
    and \citet{curtis_temporary_2019}, respectively.
    \label{fig:rotation}
	}
\end{figure}

Figure~\ref{fig:rotation} compares the stellar rotation period of \tn\ to
periods that have been measured in a number of well-studied open clusters.
\tn\ seems to be gyrochronologically coeval with the Pleiades sequence.

\begin{figure}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Lithium 6708\AA\ equivalent widths for \tn, field stars, and
    young open clusters.}
    The field star sample is drawn from Kepler planet hosts,
    and was measured by
    \citet{berger_identifying_2018} using Keck-HIRES.
    The young open clusters members were surveyed by
    \citet{randich_gaiaeso_2018} using the UVES and GIRAFFE
    spectrographs at the ESO VLT.  \citet{randich_gaiaeso_2018}
    found lithium depletion boundary ages for these clusters of
    $37.7^{+5.7}_{-4.8}\,{\rm Myr}$ (NGC$\,$2547) and
    $43.7^{+4.3}_{-3.9}\,{\rm Myr}$ (\cn).
    \label{fig:lithium}
	}
\end{figure}

Figure~\ref{fig:lithium} compares our measured \tn\ Li EW 
to stars in the field, and to young open cluster members.
The field star measurements were collected by
\citet{berger_identifying_2018}.
We have shown the points with lithium detections at ${\rm SNR}>3$.
TOI-837 is consistent with being younger than almost all CKS
planet-hosting field stars.

\subsection{Stellar Parameters}

\subsection{RVs}
I don't strongly expect the Veloce RVs or the FEROS ones to lead to a
mass measurement. The reason is that with vsini=17km/s, and 2\%
rotation amplitude signal, we expect an RV RMS of 300m/s at the
Prot=3.5 day rotation period (vsini*rotation amplitude). This is
probably larger than the expected planet signal (100m/s, 8 days).

\section{Validation of \pn}
\label{subsec:validation}

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f10.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Astrophysical false positive scenarios.}
		{\it Top}: for bound companions,
		{\it Bottom}: for unassociated companions along the same line of
		sight.
		\label{fig:fpscenario}
	}
\end{figure*}

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f11.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Multicolor photometry from the El Sauce 0.36m.}
		\label{fig:multicolor}
	}
\end{figure*}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.45\textwidth]{f3.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Scene used for blend analysis.}
		{\it Top:} Mean TESS image of \tn\ over Sector~10, with a
		logarithmic grayscale. The yellow star is the position of \tn.
		Orange circles are neighboring stars with $T<16$, scaled such that
		brighter stars are larger. The \texttt{X} and \texttt{/} hatches
		show the apertures used to measure the background and target star
		flux, respectively. Dashed lines of constant declination and right
		ascension are shown.  {\it Bottom:} Digitized Sky Survey $R$-band
		image of the same field, with a linear grayscale. The circles show
		apertures of radii 1, 1.5, and 2.25 pixels used in our blend
		analysis (Section~\ref{subsec:validation}).  Two stars of interest
		are ``Star A'' and ``Star B'', which were eventually excluded as
		being possible sources of the transits.
		\label{fig:scene}
	}
\end{figure}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.45\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ 
		{\bf SOAR HRCam contrast limits derived from point-source
			injection-recovery experiments.} 
		Star A ($\Delta T=4.7$, $\rho=2.3$'' west) is detected in the
		autocorrelation function, in addition to being a resolved Gaia
		source.
		It is co-moving with \tn, and
		its parallax and on-sky position imply that
		it is physically separated from \tn\ by $6.6\pm 0.1\,{\rm pc}$.
		\label{fig:soar}
	}
\end{figure}

@@

\section{The Planet}
\label{sec:planet}

\subsection{Transit Fitting}
Simultaneous dip plus rotation period (GP) fit. Use celerite plus PyMC3/exoplanet.
Mention FPP?

\subsection{Additional Transits}

@@

In context,...

\section{Conclusions}
\label{sec:conclusions}

>>>>

\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA {\it Transiting Exoplanet Survey Satellite} (TESS), the ESA
  Gaia mission, ground-based photometry, and spectroscopy from
  CTIO1.5/CHIRON, MPG2.2/FEROS, and AAT/Veloce.  We find that
  TOI$\,$837 is a $T=9.9$ G-dwarf in IC~2602 (the ``Southern
  Pleiades'').  The star and validated planet are therefore 30$\pm$20
  million years old.  The planet is warm ($P = 8.YY\,{\rm d}$) and
  roughly Jupiter-sized.  However, its transits are grazing: $b >
  0.92$ at 3$\sigma$.  From TESS photometry alone, the planetary size
  lies within 0.58--3.87 $R_{\rm Jup}$, due to the degeneracy between
  the planetary size and the impact parameter of the transit.  From
  radial velocity monitoring, we limit the mass of planet to less than
  2.XX $M_{\rm Jup}$ (3$\sigma$).  Grazing transits are cause for
  concern, as they usually are indicative of astrophysical false
  positive scenarios.  Our follow-up data render such scenarios highly
  unlikely.  Multi-color photometry requires eclipsing companions to
  have nearly identical color as TOI$\,$837, eliminating hierarchical
  eclipsing binary scenarios.  Background eclipsing binary scenarios
  are limited by speckle imaging, but formally remain a 0.2\%
  possibility.  TOI 837.01 is therefore a validated adolescent exoplanet
  orbiting a bright host star.
  Further analysis of its mass, atmosphere, and obliquity should
  confirm its planetary nature beyond doubt, and are worthwhile
  because young planets are expected to do interesting things.
\end{abstract}

@@

Section~\ref{sec:observations} describes the identification of the
candidate, and the follow-up observations collected.
Section~\ref{sec:validation} combines these data to assess the 
system's false positive probability, and finds that \pn\ is almost
certainly a planet.  Section~\ref{sec:system} presents our knowledge
of the cluster (Section~\ref{subsec:cluster}), the star (Section~\ref{subsec:star})
and the planet (Section~\ref{subsec:planet}).  We conclude
with a discussion in
Section~\ref{sec:discussion}.

\section{Identification and Follow-up Observations}
\label{sec:observations}

@@

%TODO: the scale of this plot is off. (Different font-sizes from
%subplots).
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f2a.pdf}
		}

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Joint TESS and multicolor photometry.}
    {\it Top:} Phase-folded TESS transit. Gray points are 2-minute
    PDCSAP flux measurements, detrended with a 0.3-day robust Huber
    spline (see Section~\ref{subsec:tess}).  Black points are binned
    to 6 minute intervals.  The line shows the transit model
    corresponding to the median transit parameters (Table~2).
    {\it Bottom:} Multicolor photometry from the El Sauce 0.36m.
    The consistent transit depths between TESS-band and the bluer
    Cousins-R and Johnson-B bandpasses rules out select false positive
    scenarios.
		\label{fig:jointphot}
	}
\end{figure*}

\begin{figure}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.47\textwidth]{f3.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Radial velocities.}
		\label{fig:rvs}
	}
\end{figure}

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Astrophysical false positive scenarios.}
		{\it Top}: for bound companions,
		{\it Bottom}: for unassociated companions along the same line of
		sight.
		\label{fig:fpscenario}
	}
\end{figure*}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.45\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Scene used for blend analysis.}
		{\it Top:} Mean TESS image of \tn\ over Sector~10, with a
		logarithmic grayscale. The yellow star is the position of \tn.
		Orange circles are neighboring stars with $T<16$, scaled such that
		brighter stars are larger. The \texttt{X} and \texttt{/} hatches
		show the apertures used to measure the background and target star
		flux, respectively. Dashed lines of constant declination and right
		ascension are shown.  {\it Bottom:} Digitized Sky Survey $R$-band
		image of the same field, with a linear grayscale. The circles show
		apertures of radii 1, 1.5, and 2.25 pixels used in our blend
		analysis (Section~\ref{subsec:validation}).  Two stars of interest
		are ``Star A'' and ``Star B'', which were eventually excluded as
		being possible sources of the transits.
		\label{fig:scene}
	}
\end{figure}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.45\textwidth]{f6.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ 
		{\bf SOAR HRCam contrast limits derived from point-source
			injection-recovery experiments.} 
		Star A ($\Delta T=4.7$, $\rho=2.3$'' west) is detected in the
		autocorrelation function, in addition to being a resolved Gaia
		source.
		It is co-moving with \tn, and
		its parallax and on-sky position imply that
		it is physically separated from \tn\ by $6.6\pm 0.1\,{\rm pc}$.
		\label{fig:soar}
	}
\end{figure}

\begin{figure}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.45\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ 
  {\bf Hertzsprung-Russell diagram of \tn\ and members of \cn.}
  Members (black circles) were identified by
  \citet{cantatgaudin_gaia_2018}.  Gray circles are non-member stars
  within 5 standard deviations of the mean \cn\ right ascension,
  declination, and parallax.  $G$ denotes Gaia broadband magnitudes,
  $Bp$ Gaia blue, $Rp$ Gaia red, and $\omega_{\rm as}$ the parallax in
  arcseconds. 
  \label{fig:hr}
	}
\end{figure}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.45\textwidth]{f9a.pdf}
		}

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.45\textwidth]{f9b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Youth diagnostics.}
    {\it Top:} Lithium 6708\AA\ equivalent widths for \tn, field
    stars, and young open clusters.  The field star sample is drawn
    from Kepler planet hosts, and was measured by
    \citet{berger_identifying_2018} using Keck-HIRES.  The young open
    clusters members were surveyed by \citet{randich_gaiaeso_2018}
    using the UVES and GIRAFFE spectrographs at the ESO VLT.
    \citet{randich_gaiaeso_2018} found lithium depletion boundary ages
    for these clusters of $37.7^{+5.7}_{-4.8}\,{\rm Myr}$
    (NGC$\,$2547) and $43.7^{+4.3}_{-3.9}\,{\rm Myr}$ (\cn).  {\it
    Bottom:} Rotation periods for \tn\ and selected open clusters.
    The Pleiades (120$\,$Myr), Praesepe (670$\,$Myr), and NGC$\,$6811
    (1000$\,$Myr) are shown.  Their rotation periods were measured by
    \citet{rebull_rotation_2016a,douglas_poking_2017,douglas_k2_2019},
    and \citet{curtis_temporary_2019}, respectively.
    \label{fig:lithium_rotation}
	}
\end{figure}

@@

\section{Elimination of False Positive Scenarios}
\label{sec:validation}

@@

\section{System Modelling}
\label{sec:system}

\subsection{The Star}
\label{subsec:star}
\subsubsection{Cluster Membership}
\label{subsec:member}

\subsubsection{HR Diagram}

Figure~\ref{fig:lithium_rotation} compares the stellar rotation period of \tn\ to
periods that have been measured in a number of well-studied open clusters.
\tn\ seems to be gyrochronologically coeval with the Pleiades sequence.

Figure~\ref{fig:lithium_rotation} compares our measured \tn\ Li EW 
to stars in the field, and to young open cluster members.
The field star measurements were collected by
\citet{berger_identifying_2018}.
We have shown the points with lithium detections at ${\rm SNR}>3$.
TOI-837 is consistent with being younger than almost all CKS
planet-hosting field stars.

\subsubsection{Stellar Parameters}

\subsubsection{RVs}
I don't strongly expect the Veloce RVs or the FEROS ones to lead to a
mass measurement. The reason is that with vsini=17km/s, and 2\%
rotation amplitude signal, we expect an RV RMS of 300m/s at the
Prot=3.5 day rotation period (vsini*rotation amplitude). This is
probably larger than the expected planet signal (100m/s, 8 days).

\subsection{The Planet}
\label{subsec:planet}

\subsubsection{Transit Fitting}
Simultaneous dip plus rotation period (GP) fit. Use celerite plus PyMC3/exoplanet.
Mention FPP?

\subsubsection{Additional Transits}

@@

Rossiter measurements will confirm.

After, a mass measurement should be worthwhile.

We'll get more TESS data in January(?).
